---
title: "Figure 2: Core Microbiome Analysis"
output: html_notebook
---

Recommendation for reading and cutoffs:
https://www.pnas.org/doi/10.1073/pnas.2104429118


Overview:
1. PCoA Analysis - complete microbiome using Unifrac? distances
2. Alpha diversity and betadispersion 
3. Determine the core across control samples 
4. How much of the community is captured?
5. How does community change with exposure to ABS? 
6. Which core members are differential abundant at TP4,TP5,TP6? 

```{r}
library(tidyr)
library(tidyverse)
library(ggplot2)
library(vegan)
library(phyloseq)
library(ggrepel)
library(patchwork)
library(microbiome)
library(RColorBrewer)
```

plotting conditions
```{r}
trt16<-c('ABS1','ABS2','Control','Experimental', 'FASW','Priming','blank', 'FAASW' )
l2p16<-c('indianred1','deeppink4','gray16','burlywood','burlywood','burlywood', 'gray81', 'darkgreen')
names(l2p16) = trt16

shpsc<-c(0,1,3,4,8,15,16)
names(shpsc)=c('T0','T1','T2','T3','T4','T5','T6')

theme_new<- theme_bw() + theme(panel.grid.major = element_line(color='gray90'), panel.grid.minor = element_line(color='gray90'), text= element_text(size = 15), axis.text.x = element_text(angle = 65,hjust = 1), legend.text = element_text(size = 20), axis.title = element_text(size = 15), title = element_text(size = 20))
```


read in cleaned data
```{r}
ps<-readRDS(file = 'ps.cleantree.noblanks.rds')
ps<-prune_samples(!(sample_names(ps) %in% c('EB1_S94', 'EB1_S95','EB2_S95')),ps) #remove EB samples 
ps
```
# Fig 2A: Alpha Diveristy 
```{r}
ps.red<-subset_samples(ps, Time %in% c('T0','T1','T4','T5','T6'))

p<-plot_richness(ps.red, measures=c("Observed"))
p_df<-p$data

a<-ggplot(p_df, aes(y=value, x=Condition, fill=Condition)) + 
  geom_boxplot(alpha=0.5, outlier.shape = NA, color='black') + 
   scale_fill_manual(values=l2p16)+scale_color_manual(values=l2p16)+
  geom_jitter(width=0.1,height=0,size=1, alpha=0.5) + facet_grid(~Time,scales='free',space='free')+ 
  theme_new + ylab('Nr. ASVs') + xlab('') + guides(fill=F, group=F,color=F) + ylim(c(0,140)) + 
  theme(axis.text.x = element_blank(),panel.grid = element_blank(),panel.grid.minor = element_blank(),rect  = element_blank())
a
```
stats
```{r}
p_df.trt<-p_df 
hist(p_df.trt$value)
p_df.trt$Time<-as.factor(p_df.trt$Time)
p_df.trt$Condition<-as.factor(p_df.trt$Condition)


aov.trt<-aov(value ~ Condition*Time, p_df.trt)
shapiro.test(residuals(aov.trt))
summary(aov.trt)
aov.trt
par(mfrow=(c(2,2)))
plot(aov.trt)
```
```{r}
data_summary <- group_by(p_df.trt, Condition, Time) %>%
  summarise(mean=mean(value), sd=sd(value), se=sd(value)/sqrt(length(value))) %>%
  arrange(desc(mean))
print(data_summary)
```

```{r}
tukey <- TukeyHSD(aov.trt)
print(tukey)
```
```{r}
head(p_df)
p_df %>% group_by(Time, Condition) %>% summarize(mean_asvs=mean(value), se_asv=sd(value)/sqrt(length(value)))
```
```{r}
#replace NAs with 0s to get multcomp to run 
tukey$`Condition:Time`[!complete.cases(tukey$`Condition:Time`),] <- 0
```

```{r}
library(multcompView)
tukey.cld <-multcompLetters4(aov.trt, tukey)
print(tukey.cld)
```
```{r}
tkl<-as.vector(tukey.cld$`Condition:Time`)
tkl.df<-data.frame(trt=tkl['Letters'])
tkl.df<-tkl.df %>% rownames_to_column() %>% separate(rowname,into=c('Condition','Time'))
head(tkl.df)
tkl.df$Condition<-factor(tkl.df$Condition, ordered=T, levels=c('Control','Experimental','ABS1','ABS2'))
tkl.df<-tkl.df[!tkl.df$Letters %in% c('d','e','f','g','h','i','j','k'),]
```

```{r}
head(p_df)
p_df$Condition<-factor(p_df$Condition, ordered=T, levels=c('Control','Experimental','ABS1','ABS2'))

a<-ggplot(p_df, aes(y=value, x=Condition, fill=Condition)) + 
  geom_boxplot(alpha=0.5, outlier.shape = NA, color='black') + 
  geom_text(data=tkl.df, aes(y=10,x=Condition,label=Letters))+
   scale_fill_manual(values=l2p16)+scale_color_manual(values=l2p16)+
  geom_jitter(width=0.1,height=0,size=1, alpha=0.5) + facet_grid(~Time,scales='free',space='free')+ 
  theme_new + ylab('Nr. ASVs') + xlab('') + guides(fill=F, group=F,color=F) + ylim(c(0,140)) + 
  theme(axis.text.x = element_blank(),panel.grid = element_blank(),panel.grid.minor = element_blank(),rect  = element_blank())
a

```



# Indentify the Core Analyisis
ASVs present in at least 90% of control samples at any relative abundance
```{r}
# subset for control group
ps.control<-subset_samples(ps.red, Condition %in% c('Control'))

#removes taxa with 0 counts
ps.control <- prune_taxa(taxa_sums(ps.control) > 0, ps.control)

#transforms to rel. abundance 
pseq.rel <- microbiome::transform(ps.control, "compositional")
```

plots across prevalance thresholds to help guide selection of the core
```{r}
# Core with compositionals:
prevalences <- seq(.05, 1, .05)
detections <- round(10^seq(log10(1e-3), log10(.2), length = 10), 3)

# Also define gray color palette
gray <- gray(seq(0,1,length=5))
# Only include core taxa
pseq.core <- core(pseq.rel, detection =0, prevalence = 90/100)
p <- plot_core(pseq.core, plot.type = "heatmap",
           colours = rev(brewer.pal(5, "Spectral")),           
               prevalences = prevalences,
           detections = detections
           ) +
    labs(x = "Detection Threshold") 
print(p)  
```
Subset entire relative abundance dataset with only core taxa
```{r}
##identify core taxa
core.taxa <- core_members(pseq.rel, detection = 0, prevalence = 90/100)
ps.red.rel<-microbiome::transform(ps.red, "compositional")
ps.core<-subset_taxa(ps.red.rel, rownames(tax_table(ps.red.rel)) %in% core.taxa) #subsets relative abundance table from the ps.red sample dataset 
ps.core
```
51 Taxa make up the Aiptasia "core" as defined by prevalence in 90 % of all control samples across time points at any relative abundance 

```{r}
range(sample_sums(ps.core))
range(sample_sums(subset_samples(ps.core, Condition=='Experimental')))
range(sample_sums(subset_samples(ps.core, Condition=='ABS1')))
range(sample_sums(subset_samples(ps.core, Condition=='ABS2')))
```
```{r}
microbiome::plot_composition(ps.core, 
    plot.type = "barplot", sample.sort = "neatmap",group_by = 'Condition') 
```

```{r}
core.taxa.df<-data.frame(tax_table(ps.core)) %>% arrange(Phylum, Class)
core.taxa.df
```
```{r}
core.taxa.df %>% group_by(Class) %>% summarize(nr=n())
```


# Fig 2B & C - betadiversity 

```{r}
# ordinate
set.seed(99)
ord<-ordinate(ps.core,"PCoA",distance = "bray")
ord.df<-plot_ordination(ps.core,ord,justDF=T)

fb<-ggplot(ord.df, aes(x=Axis.1, y =Axis.2, color=Condition, shape = Time, Fill=Condition)) + 
  geom_point(size=4) + 
  scale_shape_manual(values=shpsc) + 
  scale_color_manual(values=l2p16) + scale_fill_manual(values=l2p16) +
  theme_new #+ guides(shape=F, color=F)
fb
```
betadispersion
```{r}
#run adonis
iDist <- phyloseq::distance(ps.core, method = "bray")  # note, must call phyloseq specifically if deseq2 is also loaded
pn_df <- data.frame(sample_data(ps.core))

#betadispersion
samps<-data.frame(sample_data(ps.red))

#betdisper
bd<-betadisper(iDist, paste(samps$Time,samps$Condition))
bd.df<-data.frame(distances=bd$distances, group1=bd$group,group2=bd$group)
bd.df<-bd.df %>% separate(group1,into = c('Time','Treatment'))
bd.df$Treatment<-factor(bd.df$Treatment, ordered=T, levels=c('Control','Experimental','ABS1','ABS2'))

c<-ggplot(bd.df, aes(group=group2, y=distances, x=Treatment, fill=Treatment)) + 
  geom_boxplot(alpha=0.5, outlier.shape = NA,color='black') + 
  scale_fill_manual(values=l2p16)+ scale_color_manual(values=l2p16)+
  geom_jitter(width=0.1,height=0,size=1, alpha=0.5) + facet_grid(~Time,scales='free',space='free')+ 
  theme_new + ylab('Distance to centriod') + xlab('') + guides(fill=F, group=F, color=F) + 
  theme(axis.text.x = element_blank(),panel.grid = element_blank(),plot.background = element_blank(),panel.grid.minor = element_blank(),rect  = element_blank())
c
```

stats
```{r}
# Adonis test
mod<-adonis2(iDist ~ Condition*Time, data = pn_df)
mod
plot(mod)

#betadispersion
aov<-anova(bd)
aov
bd
```
```{r}
tukey.bd<-TukeyHSD(aov, ordered=T)
tukey.bd
```

```{r}
#replace NAs with 0s to get multcomp to run 
tukey$`Condition:Time`[!complete.cases(tukey$`Condition:Time`),] <- 0
```

```{r}
library(multcompView)
tukey.cld <-multcompLetters4(aov,comp = tukey.bd)
print(tukey.cld)
multcompLetters(tukey.bd$group[,4])
```


#Fig 2D
plot distance tree 
```{r}
tr<-plot_tree(ps.core, 'treeonly',label.tips="taxa_names",nodelabf=nodeplotblank,ladderize="left", justify = "yes please")
asv.order<-data.frame(tr$data) %>% arrange(y) 
asv.order<-na.omit(as.vector(asv.order$OTU))
asv.order
```
```{r}
d<-ggtree::ggtree(ps.core) + ggtree::geom_tiplab(align=True, hjust=-.3) + ggtree::hexpand(.01) 
d
ggtree::ggtree(ps.core) + ggtree::geom_tiplab(align=F, hjust=-.3, aes(label=Family)) + ggtree::hexpand(.01) 
```
```{r}
library(ggtree)
ggtree(ps.core) + geom_tiplab(align=F, hjust=-.3, aes(label=Class))
```

# Fig 2E

```{r}
# Bubble plot of relative abundances in ps.core
# Include only 1 control sample per time point to reduce complexity of plot 
ps.core.control<-subset_samples(ps.core,Condition=='Control') %>%  subset_samples(Sample_names %in%c('SAM_133_S1','SAM_157_S10','SAM_347_S56','SAM_364_S73','SAM_409_S89'))
ps.core.other<-subset_samples(ps.core,!Condition=='Control') 

#merge
ps.core.new<-merge_phyloseq(ps.core.control,ps.core.other)

# place samples into correct order
control.samps<-data.frame(sample_data(ps.core.control))$Sample_names
other.samps<-data.frame(sample_data(ps.core.other))$Sample_names
samps.order<-c(control.samps,other.samps)

#Add Grouping variable for plot 
samps.new<-data.frame(sample_data(ps.core.new)) %>% mutate(PlotGroup=case_when(Condition=='Control' ~ 'ASW', 
                                                                               Condition=='Experimental' ~ 'FASW 33 d',
                                                                               Condition=='ABS1' & Time=='T4' ~ "ABS1 55 d",
                                                                                Condition=='ABS1' & Time=='T5' ~ "ABS1 61 d",
                                                                                Condition=='ABS1' & Time=='T6' ~ "ABS1 76 d",
                                                                                Condition=='ABS2' & Time=='T4' ~ "ABS2 55 d",
                                                                                Condition=='ABS2' & Time=='T5' ~ "ABS2 61 d",
                                                                                Condition=='ABS2' & Time=='T6' ~ "ABS2 76 d"))

sample_data(ps.core.new)<-samps.new
```


```{r}
#make color palette for core taxa via family level classification 
df.taxa<-data.frame(ASV=rownames(tax_table(ps.core.new)),tax_table(ps.core.new)) %>% arrange(Phylum, Class, Order, Family, Genus)
df.taxa<-data.frame(ASV=rownames(tax_table(ps.core.new)),tax_table(ps.core.new)) %>% arrange(factor(ASV, levels=asv.order))
df.taxa$Family2<-ifelse(is.na(df.taxa$Family),paste(df.taxa$Class,'-unknown',sep=''),df.taxa$Family) #fix na family level classificaiton 
df.taxa$Family2<-ifelse(df.taxa$Family2=='NA-unknown',paste(df.taxa$Phylum,'-unknown',sep=''),df.taxa$Family2) #fix na family level classificaiton 
df.taxa

#set color pallet based on taxa order in tree, reds gammas, yellows, alphas, purples 
clr<-c('#FCD9DF','#38050D','#F7A1AF','#F05670','#EB1E40','#CE1231','#960D24', #"Alteromonadaceae","Gammaproteobacteria-unknown","Vibrionaceae","Halomonadaceae","Pseudomonadaceae","Oleiphilaceae",Spongiibacteraceae 
  '#D46021',# Akkermansiaceae
  'mediumpurple4','mediumpurple1', #"Phycisphaerae-unknown","Pirellulaceae" 
   '#9BC0AF', #Nannocystaceae
  '#187B4D', #Mycobacteriaceae
  '#093149','#49AEE9','#A4D7F4',#"Bacteroidia-unknown","Saprospiraceae","Flavobacteriaceae" 
  '#F9F5DC',"#F0E7A8",'#F7EB8D',"#E7DA73",#"Emcibacteraceae","Hyphomonadaceae","GCA-2696645","Sphingomonadaceae" 
  "gray",#"Proteobacteria-unknown"
  '#F3E04B','#574F0F','#F0DA32','#E4CB11','#BEA90E','#85760A'#"Devosiaceae","Stappiaceae","Alphaproteobacteria-unknown", "Cohaesibacteraceae"
  #"Rhodobacteraceae","Rhizobiaceae" 
)    


names(clr)<-unique(df.taxa$Family2)
```

make bubble plot 
```{r}
# get relavent dataframes 
md<-data.frame(sample_data(ps.core.new))
df.asv<-data.frame(Sample_names=rownames(otu_table(ps.core.new)),otu_table(ps.core.new))

#make long dataframe for plotting
df.asv.long<-pivot_longer(df.asv,cols=!Sample_names,names_to='ASV',values_to = 'RelAbund') #melt dataframe
df.relabund<-df.asv.long %>% left_join(y=md, by='Sample_names') #add in metadata
df.relabund<-df.relabund %>% left_join(y=df.taxa, by='ASV') #add in taxa info
df.relabund<-df.relabund[df.relabund$RelAbund>0,]

#re-order ASVs
df.relabund$ASV<-factor(df.relabund$ASV, ordered=T, levels=asv.order)

#reoder samples 
df.relabund$Sample_names<-factor(df.relabund$Sample_names, ordered=T, levels=samps.order)

#re-order treatments
unique(df.relabund$PlotGroup)
df.relabund$PlotGroup<-factor(df.relabund$PlotGroup, ordered=T, levels=c("ASW","FASW 33 d", "ABS1 55 d", "ABS2 55 d", "ABS1 61 d", "ABS2 61 d", "ABS1 76 d","ABS2 76 d"))

head(df.relabund)
e<-ggplot(df.relabund, aes(x = Sample_names, y = ASV, fill = Family2)) +
  geom_point(aes(size = RelAbund*10), alpha = 0.8, shape = 21) + 
  scale_size_continuous(limits = c(0, 70), range = c(0,20), breaks = c(0.01, 5, 10, 20, 40, 80)) +
  theme_new + xlab('') +ylab('') + 
  facet_grid(~PlotGroup, scales='free',space='free')+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(size = "Relative Abundance", color = "Family")  + 
  scale_fill_manual(values=clr) + guides(fill=F)
e
#ggsave('maggie_plots/bubbles-with-legend.pdf')
```
put all together 
```{r}
library(patchwork)

layout3<-"AABBBB
CCBBBB
DEEEEE
DEEEEE
DEEEEE
"

e2<-e+theme(panel.grid = element_blank(),rect = element_blank(),panel.grid.minor = element_blank(),panel.grid.major = element_blank())
e2

fb2<-fb + guides(color="none", fill="none", shape="none")

a + fb2 + c + d + e2 + plot_layout(design=layout3)

ggsave('Figures/Figure2-new.pdf', width=11,height=10)
```



```{r}
#save core taxa as .csv
df.taxa<-df.taxa %>% unite('string',c(Phylum,Class,Order,Family,Genus,Species), sep='_;',na.rm=F, remove=F)

head(df.taxa)
write.csv(df.taxa, 'Figures/core.taxa.csv')
```


Figure S1
```{r}
samps<-data.frame(sample_data(ps.core))
samps$Condition<- factor(samps$Condition, ordered=T, levels=c('Control', 'Experimental','ABS1','ABS2'))
samps<-samps %>% arrange(Condition,Time)

#make new grouping factor 
samps$grp<-paste(samps$Condition, samps$Time)
samps$grp<-factor(samps$grp, ordered = T, levels=c("Control T0","Control T1","Control T4", "Control T5","Control T6","Experimental T1", "ABS1 T4","ABS1 T5","ABS1 T6","ABS2 T4", "ABS2 T5","ABS2 T6"))
sample_data(ps.core)<-samps

ps.core %>%
  aggregate_taxa(level='Family') %>% 
  plot_composition(plot.type='barplot',group_by = 'grp',sample.sort = 'neatmap')  +
  scale_fill_manual(values=clr)+
  scale_y_continuous(label = scales::percent) + 
  theme_new + guides(fill=F)+
  # Removes sample names and ticks
  theme( axis.text.x = element_blank(), axis.ticks.x=element_blank(), plot.background = element_blank()) 
ggsave('Figures/figS1.pdf',height=6,width=12)
```


