---
title: "Figure 3: Differential Abundance Analysis"
output:
  html_document:
    df_print: paged
---
Overview- DA for Macvittie et. al Microbiome Manipulation including: 
1. Compare Diff. Abundance between Control and ABS1
2. Compare Diff. Abundance between Control and ABS2
3. Figure3 plotting Diff. Expressed taxa

```{r}
library(tidyr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(patchwork)
library(dplyr)
library(lubridate)
library(vegan)
library(ggpubr)
library(rstatix)
library(stringr)
library(emmeans)
library(multcomp)
library(multcompView)
library(dada2)
library(phyloseq)
library(ggh4x)
library(grid)
library(ALDEx2)
library(ggrepel)
```

#Read in Phyloseq object for 
```{r}
ps<-readRDS(file = 'rds_objects/ps.cleantree.noblanks.rds') 
```

# 1. ABS1 vs. Control 

subset samples to compare only abs1 and control at each timepoint 
```{r}
ps_abs1 <- subset_samples(ps, Condition == "Control"| Condition == "ABS1")
ps_abs1_t4 <-subset_samples(ps_abs1, Time == "T4") #  abs exposure
ps_abs1_t5 <-subset_samples(ps_abs1, Time == "T5") # recovery point 1
ps_abs1_t6 <-subset_samples(ps_abs1, Time == "T6") # recovery point 2
```

### ABS1 T4 - end of ABS exposure 
```{r}
sample_data(ps_abs1_t4) #checking if it looks like i want it to look like 
```

run aldex2
```{r}
pf<-ps_abs1_t4 #subset phyloseq object here, you can copy and paste the next few sections to easily repeat - just change output and input names
conds <- sample_data(pf)$Condition
reads<-t(data.frame(otu_table(pf)))
reads<-reads[rowSums(reads) > 0,] #removes rows with only 0 in dataframe

#run clr
x <- aldex.clr(reads, conds, mc.samples=128, denom="all", verbose=F)

#run t-test because we are only comparing one variable with two levels
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.kw <- aldex.kw(x)

#run effect
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

#combine table
x.all <- data.frame(x.tt,x.kw,x.effect)
# Plot 
par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="glm", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="glm",cutoff.pval=0.05)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="welch",cutoff.pval=0.05)

# Check for asymmetry - should be centered around 0
par(mfrow=c(1,2))
hist(x.all$diff.btw)
hist(x.all$effect)
```


plots from aldex package
```{r}
par(mfrow=c(1,2))
plot(x.all$effect, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab=" BH P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
abline(v=2, lty=2, col="grey")
abline(v=-2, lty=2, col="grey")

plot(x.all$diff.btw, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
```

subset ASVs with an effect size > 2
pvalue for glm < 0.05 and ttest pvalue < 0.05
nr. sigtaxa
```{r}
x.all.abs1.t4<-x.all
x.all.sig.abs1.t4<-x.all.abs1.t4 %>% filter(effect > 2 | effect < -2) %>% filter(glm.eBH < 0.05) %>% filter(we.eBH < 0.05)
x.all.sig.abs1.t4
sig_features_abs1_t4<-rownames(x.all.sig.abs1.t4)
length(sig_features_abs1_t4)
```
volcano plot for supplement
```{r}
ggplot(x.all.abs1.t4, aes(y=-log10(we.eBH), x=effect)) + 
  geom_point() + 
  geom_point(data=x.all.sig.abs1.t4, aes(y=-log10(we.eBH), x=effect),color='blue') + 
  geom_hline(yintercept = -log10(0.05),color='gray', linetype='dashed') + geom_vline(xintercept = 2, color='gray', linetype='dashed') +  geom_vline(xintercept = -2,color='gray', linetype='dashed')+
  geom_text_repel(data=x.all.sig.abs1.t4,aes(y=-log10(we.eBH), x=effect,label=rownames(x.all.sig.abs1.t4)), min.segment.length=0, max.overlaps = 15) + 
  theme_bw() + ggtitle(label = 'ABS1 vs. Control')
```

### ABS1 - T5
```{r}
pf<-ps_abs1_t5 #subset phyloseq object here, you can copy and paste the next few sections to easily repeat - just change output and input names
conds <- sample_data(pf)$Condition
reads<-t(data.frame(otu_table(pf)))
reads<-reads[rowSums(reads) > 0,] #removes rows with only 0 in dataframe

#run clr
x <- aldex.clr(reads, conds, mc.samples=128, denom="zero", verbose=F)

#run t-test because we are only comparing one variable with two levels
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.kw <- aldex.kw(x)

#run effect
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

#combine table
x.all <- data.frame(x.tt,x.kw,x.effect)

# Plot 
par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="glm", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="glm",cutoff.pval=0.05)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="welch",cutoff.pval=0.05)

# Check for asymmetry - should be centered around 0
par(mfrow=c(1,2))
hist(x.all$diff.btw)
hist(x.all$effect)
```


plots from aldex package
```{r}
par(mfrow=c(1,2))
plot(x.all$effect, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab=" BH P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
abline(v=2, lty=2, col="grey")
abline(v=-2, lty=2, col="grey")

plot(x.all$diff.btw, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
```

subset ASVs with an effect size > 2
pvalue for glm < 0.05 and ttest pvalue < 0.05
nr. sigtaxa
```{r}
x.all.abs1.t5<-x.all
x.all.sig.abs1.t5<-x.all.abs1.t5 %>% filter(effect > 2 | effect < -2) %>% filter(glm.eBH < 0.05) %>% filter(we.eBH < 0.05)
x.all.sig.abs1.t5
sig_features_abs1_t5<-rownames(x.all.sig.abs1.t5)
length(sig_features_abs1_t5)
```
volcano plot for supplement
```{r}
ggplot(x.all.abs1.t5, aes(y=-log10(we.eBH), x=effect)) + 
  geom_point() + 
  geom_point(data=x.all.sig.abs1.t5, aes(y=-log10(we.eBH), x=effect),color='blue') + 
  geom_hline(yintercept = -log10(0.05),color='gray', linetype='dashed') + geom_vline(xintercept = 2, color='gray', linetype='dashed') +  geom_vline(xintercept = -2,color='gray', linetype='dashed')+
  geom_text_repel(data=x.all.sig.abs1.t5,aes(y=-log10(we.eBH), x=effect,label=rownames(x.all.sig.abs1.t5)), min.segment.length=0, max.overlaps = 15) + 
  theme_bw() + ggtitle(label = 'ABS1 vs. Control')
```
### ABS1 - T6
```{r}
pf<-ps_abs1_t6 #subset phyloseq object here, you can copy and paste the next few sections to easily repeat - just change output and input names
conds <- sample_data(pf)$Condition
reads<-t(data.frame(otu_table(pf)))
reads<-reads[rowSums(reads) > 0,] #removes rows with only 0 in dataframe

#run clr
x <- aldex.clr(reads, conds, mc.samples=128, denom="all", verbose=F)

#run t-test because we are only comparing one variable with two levels
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.kw <- aldex.kw(x)

#run effect
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

#combine table
x.all <- data.frame(x.tt,x.kw,x.effect)

# Plot 
par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="glm", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="glm",cutoff.pval=0.05)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="welch",cutoff.pval=0.05)

# Check for asymmetry - should be centered around 0
par(mfrow=c(1,2))
hist(x.all$diff.btw)
hist(x.all$effect)
```


plots from aldex package
```{r}
par(mfrow=c(1,2))
plot(x.all$effect, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab=" BH P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
abline(v=2, lty=2, col="grey")
abline(v=-2, lty=2, col="grey")

plot(x.all$diff.btw, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
```

subset ASVs with an effect size > 2
pvalue for glm < 0.05 and ttest pvalue < 0.05
nr. sigtaxa
```{r}
x.all.abs1.t6<-x.all
x.all.sig.abs1.t6<-x.all.abs1.t6 %>% filter(effect > 2 | effect < -2) %>% filter(glm.eBH < 0.05) %>% filter(we.eBH < 0.05)
x.all.sig.abs1.t6
sig_features_abs1_t6<-rownames(x.all.sig.abs1.t6)
length(sig_features_abs1_t6)
```
volcano plot for supplement
```{r}
ggplot(x.all.abs1.t6, aes(y=-log10(we.eBH), x=effect)) + 
  geom_point() + 
  geom_point(data=x.all.sig.abs1.t6, aes(y=-log10(we.eBH), x=effect),color='blue') + 
  geom_hline(yintercept = -log10(0.05),color='gray', linetype='dashed') + geom_vline(xintercept = 2, color='gray', linetype='dashed') +  geom_vline(xintercept = -2,color='gray', linetype='dashed')+
  geom_text_repel(data=x.all.sig.abs1.t6,aes(y=-log10(we.eBH), x=effect,label=rownames(x.all.sig.abs1.t6)), min.segment.length=0, max.overlaps = 15) + 
  theme_bw() + ggtitle(label = 'ABS1 vs. Control')
```
Make new phyloseq object and save
```{r}
feature_set_abs1<-unique(c(sig_features_abs1_t6,sig_features_abs1_t5,sig_features_abs1_t4))
temp<-transform_sample_counts(ps_abs1, function(OTU) OTU/sum(OTU) * 100) #convert to rel abundance
ps_abs1.aldex.sig.ra<-prune_taxa(feature_set_abs1, temp)
```
```{r}
save(list=c('ps_abs1.aldex.sig.ra','x.all.abs1.t6','x.all.sig.abs1.t6','x.all.abs1.t5','x.all.sig.abs1.t5','x.all.abs1.t4','x.all.sig.abs1.t4'),file='rds_objects/sig.features.abs1.aldex.rds')
```

# 2. ABS2 vs Control

subset samples to compare only abs1 and control at each timepoint 
```{r}
ps_abs2 <- subset_samples(ps, Condition == "Control"| Condition == "ABS2")
ps_abs2_t4 <-subset_samples(ps_abs2, Time == "T4") #  abs exposure
ps_abs2_t5 <-subset_samples(ps_abs2, Time == "T5") # recovery point 1
ps_abs2_t6 <-subset_samples(ps_abs2, Time == "T6") # recovery point 2
```

### ABS2 T4 - end of ABS exposure 
```{r}
sample_data(ps_abs2_t4) #checking if it looks like i want it to look like 
```

run aldex2
```{r}
pf<-ps_abs2_t4 #subset phyloseq object here, you can copy and paste the next few sections to easily repeat - just change output and input names
conds <- sample_data(pf)$Condition
reads<-t(data.frame(otu_table(pf)))
reads<-reads[rowSums(reads) > 0,] #removes rows with only 0 in dataframe

#run clr
x <- aldex.clr(reads, conds, mc.samples=128, denom="all", verbose=F)

#run t-test because we are only comparing one variable with two levels
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.kw <- aldex.kw(x)

#run effect
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

#combine table
x.all <- data.frame(x.tt,x.kw,x.effect)

# Plot 
par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="glm", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="glm",cutoff.pval=0.05)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="welch",cutoff.pval=0.05)

# Check for asymmetry - should be centered around 0
par(mfrow=c(1,2))
hist(x.all$diff.btw)
hist(x.all$effect)
```


plots from aldex package
```{r}
par(mfrow=c(1,2))
plot(x.all$effect, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab=" BH P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
abline(v=2, lty=2, col="grey")
abline(v=-2, lty=2, col="grey")

plot(x.all$diff.btw, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
```
subset ASVs with an effect size > 2
pvalue for glm < 0.05 and ttest pvalue < 0.05
nr. sigtaxa
```{r}
x.all.abs2.t4<-x.all
x.all.sig.abs2.t4<-x.all.abs2.t4 %>% filter(effect > 2 | effect < -2) %>% filter(glm.eBH < 0.05) %>% filter(we.eBH < 0.05)
x.all.sig.abs2.t4
sig_features_abs2_t4<-rownames(x.all.sig.abs2.t4)
length(sig_features_abs2_t4)
```
volcano plot for supplement
```{r}
ggplot(x.all.abs2.t4, aes(y=-log10(we.eBH), x=effect)) + 
  geom_point() + 
  geom_point(data=x.all.sig.abs2.t4, aes(y=-log10(we.eBH), x=effect),color='blue') + 
  geom_hline(yintercept = -log10(0.05),color='gray', linetype='dashed') + geom_vline(xintercept = 2, color='gray', linetype='dashed') +  geom_vline(xintercept = -2,color='gray', linetype='dashed')+
  geom_text_repel(data=x.all.sig.abs2.t4,aes(y=-log10(we.eBH), x=effect,label=rownames(x.all.sig.abs2.t4)), min.segment.length=0, max.overlaps = 15) + 
  theme_bw() + ggtitle(label = 'ABS2 vs. Control')
```

### ABS2 - T5
```{r}
pf<-ps_abs2_t5 #subset phyloseq object here, you can copy and paste the next few sections to easily repeat - just change output and input names
conds <- sample_data(pf)$Condition
reads<-t(data.frame(otu_table(pf)))
reads<-reads[rowSums(reads) > 0,] #removes rows with only 0 in dataframe

#run clr
x <- aldex.clr(reads, conds, mc.samples=128, denom="all", verbose=F)

#run t-test because we are only comparing one variable with two levels
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.kw <- aldex.kw(x)

#run effect
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

#combine table
x.all <- data.frame(x.tt,x.kw,x.effect)

# Plot 
par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="glm", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="glm",cutoff.pval=0.05)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="welch",cutoff.pval=0.05)

# Check for asymmetry - should be centered around 0
par(mfrow=c(1,2))
hist(x.all$diff.btw)
hist(x.all$effect)
```


plots from aldex package
```{r}
par(mfrow=c(1,2))
plot(x.all$effect, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab=" BH P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
abline(v=2, lty=2, col="grey")
abline(v=-2, lty=2, col="grey")

plot(x.all$diff.btw, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
```

subset ASVs with an effect size > 2
pvalue for glm < 0.05 and ttest pvalue < 0.05
nr. sigtaxa
```{r}
x.all.abs2.t5<-x.all
x.all.sig.abs2.t5<-x.all.abs2.t5 %>% filter(effect > 2 | effect < -2) %>% filter(glm.eBH < 0.05) %>% filter(we.eBH < 0.05)
x.all.sig.abs2.t5
sig_features_abs2_t5<-rownames(x.all.sig.abs2.t5)
length(sig_features_abs2_t5)
```
volcano plot for supplement
```{r}
ggplot(x.all.abs2.t5, aes(y=-log10(we.eBH), x=effect)) + 
  geom_point() + 
  geom_point(data=x.all.sig.abs2.t5, aes(y=-log10(we.eBH), x=effect),color='blue') + 
  geom_hline(yintercept = -log10(0.05),color='gray', linetype='dashed') + geom_vline(xintercept = 2, color='gray', linetype='dashed') +  geom_vline(xintercept = -2,color='gray', linetype='dashed')+
  geom_text_repel(data=x.all.sig.abs2.t5,aes(y=-log10(we.eBH), x=effect,label=rownames(x.all.sig.abs2.t5)), min.segment.length=0, max.overlaps = 15) + 
  theme_bw() + ggtitle(label = 'ABS2 vs. Control')
```
### ABS2 - T6
```{r}
pf<-ps_abs2_t6 #subset phyloseq object here, you can copy and paste the next few sections to easily repeat - just change output and input names
conds <- sample_data(pf)$Condition
reads<-t(data.frame(otu_table(pf)))
reads<-reads[rowSums(reads) > 0,] #removes rows with only 0 in dataframe

#run clr
x <- aldex.clr(reads, conds, mc.samples=128, denom="all", verbose=F)

#run t-test because we are only comparing one variable with two levels
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.kw <- aldex.kw(x)

#run effect
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

#combine table
x.all <- data.frame(x.tt,x.kw,x.effect)

# Plot 
par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="glm", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="glm",cutoff.pval=0.05)

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", cutoff.pval=0.05) #can use the glm model test as the sig value 
aldex.plot(x.all, type="MW", test="welch",cutoff.pval=0.05)

# Check for asymmetry - should be centered around 0
par(mfrow=c(1,2))
hist(x.all$diff.btw)
hist(x.all$effect)
```


plots from aldex package
```{r}
par(mfrow=c(1,2))
plot(x.all$effect, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab=" BH P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
abline(v=2, lty=2, col="grey")
abline(v=-2, lty=2, col="grey")

plot(x.all$diff.btw, x.all$glm.eBH, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col='red', pch=19)
abline(h=0.05, lty=2, col="grey")
```

subset ASVs with an effect size > 2
pvalue for glm < 0.05 and ttest pvalue < 0.05
nr. sigtaxa
```{r}
x.all.abs2.t6<-x.all
x.all.sig.abs2.t6<-x.all.abs2.t6 %>% filter(effect > 2 | effect < -2) %>% filter(glm.eBH < 0.05) %>% filter(we.eBH < 0.05)
x.all.sig.abs2.t6
sig_features_abs2_t6<-rownames(x.all.sig.abs2.t6)
length(sig_features_abs2_t6)
```
volcano plot for supplement
```{r}
ggplot(x.all.abs2.t6, aes(y=-log10(we.eBH), x=effect)) + 
  geom_point() + 
  geom_point(data=x.all.sig.abs2.t6, aes(y=-log10(we.eBH), x=effect),color='blue') + 
  geom_hline(yintercept = -log10(0.05),color='gray', linetype='dashed') + geom_vline(xintercept = 2, color='gray', linetype='dashed') +  geom_vline(xintercept = -2,color='gray', linetype='dashed')+
  geom_text_repel(data=x.all.sig.abs2.t6,aes(y=-log10(we.eBH), x=effect,label=rownames(x.all.sig.abs2.t6)), min.segment.length=0, max.overlaps = 15) + 
  theme_bw() + ggtitle(label = 'ABS1 vs. Control')
```
Make new phyloseq object and save
```{r}
feature_set_abs2<-unique(c(sig_features_abs2_t6,sig_features_abs2_t5,sig_features_abs2_t4))
temp<-transform_sample_counts(ps_abs2, function(OTU) OTU/sum(OTU) * 100) #convert to rel abundance
ps_abs2.aldex.sig.ra<-prune_taxa(feature_set_abs2, temp)
```
```{r}
save(list=c('ps_abs2.aldex.sig.ra','x.all.abs2.t6','x.all.sig.abs2.t6','x.all.abs2.t5','x.all.sig.abs2.t5','x.all.abs2.t4','x.all.sig.abs2.t4'),file='./rds_objects/sig.features.abs2.aldex.rds')
```



#3. Figure 3

load relavent data
```{r}
load('./rds_objects/sig.features.abs1.aldex.rds') # contains 1 phyloseq object and all stats for each tp
load('./rds_objects/sig.features.abs2.aldex.rds') # contains 1 phyloseq object and all stats for each tp
```

set color pallet  
```{r}
taxa<-data.frame(rbind(tax_table(ps_abs1.aldex.sig.ra),tax_table(ps_abs2.aldex.sig.ra)))
taxa$ASV<-rownames(taxa)
taxa<-taxa %>% distinct(ASV, .keep_all = T) %>% arrange(Phylum,Class,Order,Family,Genus,Species)
taxa$Family2<-ifelse(is.na(taxa$Family),paste(taxa$Class,'-unknown',sep=''),taxa$Family) #fix na family level classificaiton 
taxa$Family2<-ifelse(taxa$Family2=='NA-unknown',paste(taxa$Phylum,'-unknown',sep=''),taxa$Family2) #fix na family level classificaiton 

clr<-c(
'#187B4D','#A4D7F4','#093149','mediumpurple4','mediumpurple1','#85760A', #"Mycobacteriaceae","Flavobacteriaceae","Bacteroidia-unknown","Phycisphaerae-unknown","Pirellulaceae","Rhizobiaceae"
'#BEA90E','#F9F5DC',"#E7DA73",'#F7EB8D','#FCD9DF','#F7A1AF', #"Rhodobacteraceae","Emcibacteraceae","Sphingomonadaceae","GCA-2696645","Alteromonadaceae","Vibrionaceae" 
'red','#F05670','#EB1E40','#38050D','gray'#"Francisellaceae","Halomonadaceae","Pseudomonadaceae","Gammaproteobacteria-unknown", "Proteobacteria-unknown" 
)
names(clr)<-unique(taxa$Family2)
clr

taxa<-taxa %>% unite('taxstring', Phylum,Class,Order,Family,Genus,Species,sep='_;') %>% arrange(ASV)

write.csv(taxa,'Figures/DA-taxa.csv')
```

ABS1 - Figure 3A
```{r}
# make new phyloseq object
mdata<-data.frame(sample_data(ps_abs1.aldex.sig.ra))
df.otu<-data.frame(Sample_names=rownames(otu_table(ps_abs1.aldex.sig.ra)),otu_table(ps_abs1.aldex.sig.ra))
taxa.df<-data.frame(ASV=rownames(tax_table(ps_abs1.aldex.sig.ra)),tax_table(ps_abs1.aldex.sig.ra))

df.otu.long<-pivot_longer(df.otu,cols=!Sample_names,names_to='ASV',values_to = 'RelAbund') #melt dataframe
df.relabund<-df.otu.long %>% left_join(y=mdata, by='Sample_names') #add in metadata
df.relabund<-df.relabund %>% left_join(y=taxa.df, by='ASV') #add in taxa info
df.relabund.nozeros<-df.relabund[df.relabund$RelAbund>0,]

#plot
df.abs1<-df.relabund.nozeros[df.relabund.nozeros$Time %in% c('T4','T5','T6'),]
df.abs1$Condition<-factor(df.abs1$Condition, ordered=T, levels=c('Control','ABS1'))

# Fix Unknown Families
df.abs1$Family2<-ifelse(is.na(df.abs1$Family),paste(df.abs1$Class,'-unknown',sep=''),df.abs1$Family)

#reorder asvs according to pattern
asv_order<-c('ASV151','ASV114','ASV48','ASV152','ASV68','ASV43','ASV44','ASV17', #susceptible
             'ASV3',
             'ASV5','ASV9','ASV20',
             'ASV139','ASV18', #selected for
             'ASV14','ASV24'
             )
df.abs1$ASV<-factor(df.abs1$ASV, ordered=T,levels=rev(asv_order))

```

```{r}
Fig3a<-ggplot(df.abs1, aes(x = Sample_names, y = ASV, fill = Family2)) +
  geom_point(aes(size = RelAbund*10), alpha = 0.8, shape = 21) + 
  scale_size_continuous(limits = c(0, 70), range = c(0,10), breaks = c(0.01, 5, 10, 20, 40, 80)) +
  facet_grid(. ~ Time + Condition, scales="free", space="free") +
  theme_new + xlab('Sample') +
  scale_y_discrete(limits = rev(levels(df.relabund$ASV))) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(size = "Relative Abundance", color = "Family") +
  guides(fill = guide_legend(override.aes = list(size = 2))) + 
  scale_fill_manual(values=clr) + guides(fill=F)
Fig3a
```


Fig 3b
```{r}
# make new phyloseq object
mdata<-data.frame(sample_data(ps_abs2.aldex.sig.ra))
df.otu<-data.frame(Sample_names=rownames(otu_table(ps_abs2.aldex.sig.ra)),otu_table(ps_abs2.aldex.sig.ra))
taxa.df<-data.frame(ASV=rownames(tax_table(ps_abs2.aldex.sig.ra)),tax_table(ps_abs2.aldex.sig.ra))
df.otu.long<-pivot_longer(df.otu,cols=!Sample_names,names_to='ASV',values_to = 'RelAbund') #melt dataframe
df.relabund<-df.otu.long %>% left_join(y=mdata, by='Sample_names') #add in metadata
df.relabund<-df.relabund %>% left_join(y=taxa.df, by='ASV') #add in taxa info
df.relabund.nozeros<-df.relabund[df.relabund$RelAbund>0,]

#plot
df.abs2<-df.relabund.nozeros[df.relabund.nozeros$Time %in% c('T4','T5','T6'),]
df.abs2$Condition<-factor(df.abs2$Condition, ordered=T, levels=c('Control','ABS2'))

# Fix Unknown Families
df.abs2$Family2<-ifelse(is.na(df.abs2$Family),paste(df.abs2$Class,'-unknown',sep=''),df.abs2$Family)

#order based on pattern
asv_order<-c('ASV34','ASV81','ASV68','ASV206','ASV33','ASV64','ASV101','ASV90','ASV61','ASV151','ASV114',
             'ASV104',
             'ASV39','ASV37','ASV38','ASV43','ASV26','ASV3','ASV5','ASV11','ASV19','ASV9','ASV17','ASV59',
             'ASV8','ASV20','ASV32',
             'ASV36','ASV150','ASV113')
df.abs2$ASV<-factor(df.abs2$ASV, ordered=T,levels=rev(asv_order))
```
```{r}
Fig3b<-ggplot(df.abs2, aes(x = Sample_names, y = ASV, fill = Family2)) +
  geom_point(aes(size = RelAbund*10), alpha = 0.8, shape = 21) + 
  scale_size_continuous(limits = c(0, 70), range = c(0,10), breaks = c(0.01, 5, 10, 20, 40, 80)) +
  facet_grid(. ~ Time + Condition, scales="free", space="free") +
  theme_new + xlab('Sample') +
  scale_y_discrete(limits = rev(levels(df.relabund$ASV))) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  labs(size = "Relative Abundance", color = "Family") +
  guides(fill = guide_legend(override.aes = list(size = 2))) + 
  scale_fill_manual(values=clr) + guides(fill=F)
Fig3b
```
Make figure 3
```{r}
library(patchwork)
layout<-"AA
BB
BB"
Fig3a + guides(fill=F, size=F) + Fig3b + guides(fill=F, size=F)  + plot_layout(ncol=1, design=layout)
Fig3a + Fig3b + plot_layout(ncol=1, design=layout, guides='collect')
ggsave('Figures/Figure3-withleg.pdf',width=10,height=11)
```


