---
title: "R Notebook"
output: html_notebook
---
Overview: Physiology Analysis for MacVittie et. al Microbiome Manipulation including:
1.Graphing and statistics for CFU counts
2. Graphing and statistics for Protein Biomass
3. Graphing and statistics for Pedal laceration rate
4.Graphing and statistics for Symbiont Density


# SetUp
Load packages
```{r}
library(tidyr)
library(tidyverse)
library(ggplot2)
library(vegan)
library(phyloseq)
library(ape)
library(ggh4x)
library(ggrepel)
library(patchwork)
library(ggpubr)
library(rstatix)
library(ARTool)
```

plotting conditions
```{r}
trt16<-c('ABS1','ABS1 Treatment','ABS1 Recovery','ABS2','Control','ASW','Experimental', 'FASW','Priming','blank', 'FAASW' )
l2p16<-c('indianred1','indianred1','indianred1','deeppink4','gray16','gray16','burlywood','burlywood','burlywood', 'gray81', 'darkgreen')
names(l2p16) = trt16

shpsc<-c(0,1,3,4,8,15,16)
names(shpsc)=c('T0','T1','T2','T3','T4','T5','T6')

theme_new<- theme_bw() + theme(panel.grid.major = element_line(color='gray90'), panel.grid.minor = element_line(color='gray90'), text= element_text(size = 15), axis.text.x = element_text(hjust = 1), legend.text = element_text(size = 20), axis.title = element_text(size = 15), title = element_text(size = 20),panel.border = element_blank())
theme_post<- theme_bw() + theme(panel.grid.major = element_line(color='gray90'), panel.grid.minor = element_line(color='gray90'), text= element_text(size = 15), legend.text = element_text(size = 20), axis.title = element_text(size = 15), title = element_text(size = 20),panel.border = element_blank(),panel.background = element_blank(),plot.background=element_rect(color = 'gray40',fill=NA))
```
# 1.CFU counts
```{r}
cfus<-read.csv("./data/cfu-counts-2022-01-20-modified.csv")
head(cfus)
```
## Figure1B
```{r}
Fig1B<-ggplot(cfus, aes(x=day, y=value, color=treatment)) + 
  geom_point(size=3, position=position_dodge(width = .75)) +
        scale_y_continuous(trans=scales::pseudo_log_trans(base = 10)) + 
        scale_color_manual(values=l2p16) + 
        theme_new + theme(axis.text=element_text(size=9)) +
        ylab(expression(CFU~(cells~ml^-1))) + xlab('Sampling Day') +guides(color=F)
Fig1B
```
## CFU statistics
Summary statistics
```{r}
stats<-cfus %>% group_by(day, treatment) %>%  summarise(mean_cfu=mean(value), se_cfu=sd(value)/sqrt(length(value)))
stats
```

Checking Normality and overall significance
```{r}
CFU.aov <- aov(value ~ treatment * day, cfus)
shapiro.test(residuals(CFU.aov)) # p=3.579e-14
ggdensity(cfus, x = 'value', fill = "lightgray", ) +
  stat_overlay_normal_density(color = "red", linetype = "dashed")  ## shows non-normal data distribution 
par(mfrow=c(2,2))
plot(CFU.aov)
cfus$intercfus<- interaction(cfus$treatment,cfus$day)
kruskal.test(value~intercfus, cfus) ## p = 3.941e-07
```
Violates assumptions of normality -> split data by day and run kruskal wallis test 

Kruskall Wallis test - report pvalue 
day 33- priming
```{r}
df<-cfus %>% subset(day==33)
kruskal.test(value ~ treatment, df)
(cfu_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
day 41 - treatment
```{r}
df<-cfus %>% subset(day==41)
kruskal.test(value ~ treatment, df)
(cfu_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
day 55 - treatment
```{r}
df<-cfus %>% subset(day==55)
kruskal.test(value ~ treatment, df)
(cfu_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
day 61 - treatment
```{r}
df<-cfus %>% subset(day==61)
kruskal.test(value ~ treatment, df)
(cfu_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
day 76 - treatment
```{r}
df<-cfus %>% subset(day==76)
kruskal.test(value ~ treatment, df)
(cfu_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```



# 2.Protein Biomass
```{r}
physio<-read.csv('./data/physio-2022-01-20.csv',na.strings = c('',"NA")) 
```
Data cleaning
```{r}
physio<-physio[rowSums(is.na(physio)) != ncol(physio),]
physio$expdate<-ymd(physio$date)

physio <- physio %>% mutate(Day = 
                case_when(date == "2021-11-03"~ 0,
                          date == "2021-11-05"~2,
                          date == "2021-11-12"~9,
                          date == "2021-11-18"~15,
                          date == "2021-11-29"~26,
                          date == "2021-12-06"~ 33,
                          date == "2021-12-10"~37,
                          date == "2021-12-14"~ 41,
                          date == "2021-12-17"~37,
                          date == "2021-12-21"~ 48,
                          date == "2021-12-23"~50,
                          date == "2021-12-28"~ 55,
                          date == "2022-01-04"~ 61,
                          date == "2022-01-07"~64,
                          date == "2022-01-19"~ 76))

prot<-physio[physio$variable=='avg_protein(ug/ml)',]# subsetting the protein dataset from overall physiology
```

## Fig 1C
```{r}
Fig1C<-ggplot(prot, aes(x=Day, y=value, group=interaction(timepoint,treatment), color=treatment, fill=treatment)) +
        geom_boxplot(alpha=0.5, outlier.shape = NA, position = position_dodge2(preserve = "single"))+
        geom_point(size=3, position=position_dodge(width=4)) + 
        scale_color_manual(values=l2p16,name = "Treatment") +  scale_fill_manual(values=l2p16,name = "Treatment") +
        theme_new + theme(axis.text=element_text(size=9)) +
        ylab(expression(Total~Protein~(µg~mL^-1))) + xlab("Sampling Day")+theme_new + guides(color=F, fill=F)
Fig1C
```
## Protein statistics
Summary statistics
```{r}
stats<-prot %>% group_by(Day, treatment) %>%  summarise(mean_prot=mean(value), se_prot=sd(value)/sqrt(length(value)))
stats
```

Checking Normality and overall significance
```{r}
mod<-aov(value~factor(Day)*factor(treatment), data=prot)
mod
par(mfrow=c(2,2))
plot(mod)
shapiro.test(residuals(mod)) #p=0.5335
ggdensity(prot, x = 'value', fill = "lightgray", ) +
  stat_overlay_normal_density(color = "red", linetype = "dashed")
anova(mod)
```
Running Post-Hoc tests
```{r}
tukey<-TukeyHSD(mod)
(TUK.trt<-data.frame(TukeyHSD(mod)$'factor(treatment)') %>% arrange(p.adj))
(TUK.time<-data.frame(TukeyHSD(mod)$'factor(Day)') %>% arrange(p.adj))
(TUK.txt<-data.frame(TukeyHSD(mod)$'factor(Day):factor(treatment)') %>% arrange(p.adj))
```



#3. Pedal Lacerates
subset pedal lacerates
```{r}
peds<-physio[physio$variable=='pedal_lacerates',]

peds <- peds %>% mutate(treatment = 
                case_when(description == "FASW" ~ "FASW",
                          .default = treatment))
#Remove NAs for wells where anemone died
peds <- peds[!is.na(peds$value),]
```
##Fig 1D
```{r}
# add new variable to data to break up into treatment bins 
peds <- peds %>% mutate(Phase = 
                case_when(Day == '2' ~ "Priming",
                          Day == '9' ~ "Priming",
                          Day == '15' ~ "Priming",
                          Day == '26' ~ "Priming",
                          Day == '37'~"Treatment",
                          Day == '50'~"Treatment",
                          Day == '64' ~ "Recovery 1",
                          Day == '76' ~ "Recovery 2"))

peds$Phase <- factor( peds$Phase, ordered = T, levels = c("Priming","Treatment","Recovery 1", "Recovery 2"))
#subsetting dataset to remove 2 day timepoint
peds_prime <- subset(peds, Day>2)
#graph
Fig1D<-ggplot(peds_prime, aes(x=Phase, y=value, group=interaction(Phase,treatment),color= treatment,fill=treatment)) +
        geom_boxplot(alpha=0.5, outlier.shape = NA, position = position_dodge2(preserve = "single"), width = 7.5)+
        facet_grid(~Phase, scales = "free",space = 'free',switch = 'x')+
        scale_color_manual(values=l2p16,name = "Treatment") +  scale_fill_manual(values=l2p16,name = "Treatment") +
        theme_new + theme(strip.background = element_blank(),strip.text.x = element_blank())+
        ylab(expression(Pedal~lacerates~anemone^-1~week^-1)) +guides(color=F, fill=F)
Fig1D
```

## Pedal Lacerate statistics
Summary Statistics
```{r}
#Overall
stats.peds<-peds %>% group_by(Phase,treatment) %>%  summarise(mean_peds=mean(value), se_alg=sd(value)/sqrt(length(value)))
stats.peds
#FASW pooled 
df<-peds %>% subset(Day==9 | Day==15)
stats.fasw<-df %>% group_by(Phase) %>%  summarise(mean_peds=mean(value), se_alg=sd(value)/sqrt(length(value)))
stats.fasw
```
Checking For Normality and significance
```{r}
mod<-aov(value ~ factor(Day)*factor(treatment), data=peds)
mod
par(mfrow=c(2,2))
plot(mod)
shapiro.test(residuals(mod)) #untransformed or p=2.19e-14: use kruskal wallis subset by day
ggdensity(peds, x = 'value', fill = "lightgray", ) +
  stat_overlay_normal_density(color = "red", linetype = "dashed")
#df<-peds %>% subset(Day>9), (no pedal laceration on Day 2)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj)) # use by subset of day to only compare at each timepoint
```

Violates assumptions of normality -> split data by day and run kruskal wallis test 

Kruskall Wallis test - report pvalue 
```{r}
df<-peds %>% subset(Day==9| Day==15)# cannot run 9 on its own because all values are 0
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
Day 15
```{r}
df<-peds %>% subset(Day==15)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
Day 26
```{r}
df<-peds %>% subset(Day==26)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
Day 37
```{r}
df<-peds %>% subset(Day==37)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
Day 50
```{r}
df<-peds %>% subset(Day==50)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```

Day 64
```{r}
df<-peds %>% subset(Day==64)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```
Day 76
```{r}
df<-peds %>% subset(Day==76)
kruskal.test(value ~ treatment, df)
(peds_post<-df %>% dunn_test(value ~ treatment) %>% arrange(p.adj))
```







# 4.Symbiont Density
Subsetting Data
```{r}
alg<-physio[physio$variable=='sym_prot(cells/ug)',]
alg <- alg %>% mutate(treatment = 
                case_when(description == "FASW"~ "FASW",
                          .default = treatment))
```      

## Figure 1E
```{r}
Fig1E<-ggplot(alg, aes(x=Day, y=log(value), group=interaction(Day,treatment), color=treatment, fill=treatment)) +
        geom_boxplot(alpha=0.5, outlier.shape = NA, position = position_dodge2(preserve = "single"))+
        geom_point(size=3, position=position_dodge(width=4)) + 
        scale_color_manual(values=l2p16,name = "Treatment") +  scale_fill_manual(values=l2p16,name = "Treatment") +
        theme_new + theme(axis.text=element_text(size=9)) +
        ylab(expression(Symbiont~Density~(log~cells~µg~protein^-1))) + xlab("Sampling Day")+theme(legend.position = 'top')+guides(color=F,fill=F)+theme_new 
Fig1E

Fig1_leg<- get_legend(Fig1E)# comment out guide removal in above before extracting
```
## Symbiont Statistics
Summary Statistics
```{r}
stats.alg<-alg %>% group_by(Day, treatment) %>%  summarise(mean_alg=mean(value), se_alg=sd(value)/sqrt(length(value)))
stats.alg
```

```{r}
mod.alg<-aov(log10(value)~factor(Day)*factor(treatment), data=alg)
mod.alg
par(mfrow=c(2,2))
plot(mod.alg)
alg$log.val = log10(alg$value)
shapiro.test(residuals(mod.alg)) #untransformed or p=0.0251 if log transformed p= 0.05307
ggdensity(alg, x = 'log.val', fill = "lightgray", ) +
  stat_overlay_normal_density(color = "red", linetype = "dashed")
anova(mod.alg)
```

```{r}
TukeyHSD(mod.alg)
(TUK.trt<-data.frame(TukeyHSD(mod.alg)$'factor(treatment)') %>% arrange(p.adj))
(TUK.time<-data.frame(TukeyHSD(mod.alg)$'factor(Day)') %>% arrange(p.adj))
(TUK.txt<-data.frame(TukeyHSD(mod.alg)$'factor(Day):factor(treatment)') %>% arrange(p.adj))
```
# Figure 1 Composite
```{r}
layout<-"AB
CD"

Fig1_post<-Fig1B +guides(color = F, fill=F) + Fig1C + Fig1D +guides(color=F, fill=F)+ Fig1E + plot_layout(ncol=2, design=layout, guides='collect')
Fig1  
ggsave('Figures/Fig1_post.pdf',width=8,height=8, bg= 'transparent')
ggsave('Figures/Fig1_leg.pdf',Fig1_leg,bg='transparent')
Fig1B<-Fig1B+theme_post
ggsave('Figures/Fig1B.pdf',Fig1B,bg='transparent',width = 8,height = 8)
Fig1C<-Fig1C+theme_post
ggsave('Figures/Fig1C.pdf',Fig1C,bg='transparent')
Fig1D<-Fig1D+theme_post+theme(strip.background = element_blank(),strip.text.x = element_blank())
ggsave('Figures/Fig1D.pdf',Fig1D,bg='transparent')
Fig1E<-Fig1E+theme_post
ggsave('Figures/Fig1E.pdf',Fig1E,bg='transparent')



```